---
title: "Building Model for Weight Lifting Exercise Manner Prediction"
author: "A.Volkoviy"
date: "22.11.2015"
output: html_document
---

# Building Model for Weight Lifting Exercise Manner Prediction

## Executive Summary

Weight Lifting Exercise Dataset contains data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. These participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har

This report addesses the problem of prediction the manner in which they did the exercise.

## Preparation of environment

Loading packages and seed setting

```{r}
library(caret)
library(rpart)
library(randomForest)
library(ElemStatLearn)
set.seed(2015)
```

## Loading the Data

The following R code is used for data loading:
```{r}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "pml-training.csv"
testFile  <- "pml-testing.csv"
download.file(trainUrl, destfile=trainFile, method="wget")
download.file(testUrl, destfile=testFile, method="wget")
```

Then we can read data from files:
```{r}
trainDat <- read.csv(trainFile)
testDat <- read.csv(testFile)
```

## Exploratory Analysis

Size of test set:
```{r}
dim(testDat)
```

Size of train set
```{r}
dim(trainDat)
```
And number of compleate cases are
```{r}
sum(complete.cases(trainDat))
```

Designations for exercise execution manners:
```{r}
unique(trainDat$classe, incomparables = FALSE)
```

## Preprocessing

Data should be cleaned before usage.
```{r}
trainDat <- trainDat[, colSums(is.na(trainDat)) == 0]
testDat <- testDat[, colSums(is.na(testDat)) == 0]
classe <- trainDat$classe
trainRemove <- grepl("^X|timestamp|window", names(trainDat))
trainDat <- trainDat[, !trainRemove]
trainCleaned <- trainDat[, sapply(trainDat, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testDat))
testDat <- testDat[, !testRemove]
testCleaned <- testDat[, sapply(testDat, is.numeric)]
```

Then validation set should be separated from train set:
```{r}
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=FALSE)
trainSet <- trainCleaned[inTrain, ]
validationSet <- trainCleaned[-inTrain, ]
```

## Building the model

Random Forest method is used for exercise manner prediction model 
```{r}
myModel <- train(classe ~ ., method = "rf", data = trainSet, importance = TRUE, trControl = trainControl(method = "cv", number = 5))
```

## Testing with the validation set

Prediction on validation set and confusion matrix for model performance evaluation
```{r}
validPred <- predict(myModel, newdata=validationSet)
confusionMatrix(validPred, validationSet$classe)
```
As we can see, that accuracy is 0.9929.

## Prediction

Now we can use trained model for prediction on the test set
```{r}
testingPred <- predict(myModel, newdata=testCleaned)
```

And store results to files as suggested in assignment submission instruction:
```{r}
write_files <- function(x) {
        n <- length(x)
        for (i in 1:n) {
                filename <- paste0("problem_id_", i, ".txt")
                write.table(x[i], file=filename, quote=FALSE, row.names=FALSE,col.names=FALSE)
        }
}
write_files(testingPred)
```